Using Visual Analytics to Explain\\[1ex] Black-Box Machine Learning

### 2018-04-03 aggexplain/experiment.tex
fix figures and some citations

\begin{figure*}[ht!]
\centering
\raisebox{-0.5\height}{\includegraphics[width=0.41\linewidth]{aggexplain/h_all}}
\raisebox{-0.2\height}{
\parbox{0.16\linewidth}{
\centering
\textsf{$\leftarrow$ Unbiased data set}\\
\textsf{Biased data set $\rightarrow$}\\
\textsf{Compared subset:}\\
\includegraphics[height=1em]{aggexplain/legend_all}
}
}
\raisebox{-0.5\height}{\includegraphics[width=0.41\linewidth]{aggexplain/h_all}}
~\\
\noindent\rule{\linewidth}{0.4pt}
~\\
\raisebox{-0.5\height}{\includegraphics[width=0.41\linewidth]{aggexplain/h_label_ok}}
\raisebox{-0.2\height}{
\parbox{0.16\linewidth}{
\centering
\textsf{$\leftarrow$ Unbiased data set}\\
\textsf{Biased data set $\rightarrow$}\\
\textsf{Compared subset:}\\
\includegraphics[height=1em]{aggexplain/legend_label}
}
}
\raisebox{-0.5\height}{\includegraphics[width=0.41\linewidth]{aggexplain/h_label_bad}}
~\\
\noindent\rule{\linewidth}{0.4pt}
~\\
\raisebox{-0.5\height}{\includegraphics[width=0.41\linewidth]{aggexplain/h_pred_ok}}
\raisebox{-0.2\height}{
\parbox{0.16\linewidth}{
\centering
\textsf{$\leftarrow$ Unbiased data set}\\
\textsf{Biased data set $\rightarrow$}\\
\textsf{Compared subset:}\\
\includegraphics[height=1em]{aggexplain/legend_pred}
}
}
\raisebox{-0.5\height}{\includegraphics[width=0.41\linewidth]{aggexplain/h_pred_bad}}
~\\
\noindent\rule{\linewidth}{0.4pt}
~\\
\raisebox{-0.5\height}{\includegraphics[width=0.41\linewidth]{aggexplain/h_correct_ok}}
\raisebox{-0.2\height}{
\parbox{0.16\linewidth}{
\centering
\textsf{$\leftarrow$ Unbiased data set}\\
\textsf{Biased data set $\rightarrow$}\\
\textsf{Compared subset:}\\
\includegraphics[height=1em]{aggexplain/legend_correct}
}
}
\raisebox{-0.5\height}{\includegraphics[width=0.41\linewidth]{aggexplain/h_correct_bad}}
\caption{
Comparison of different subset selections on both the unbiased model (left side) and the biased model (right side).
Features are sorted per segment by most important at the top-left, row-wise to least important at the bottom-right.
Note, the feature ``Living Area'', in both the ``Label'' (\ie, ground truth) and ``Pred.'' (\ie, model prediction), has flipped outcomes for the biased model (right side).
Each subset has a different color palette to not confuse different selections with each other.
}
\label{figs:good_vs_bad}
\end{figure*}

>>>>

\begin{figure}[t]
\centering
\raisebox{-0.5\height}{\includegraphics[width=0.4\linewidth]{aggexplain/h_all}}
\raisebox{-0.2\height}{
\parbox{0.16\linewidth}{
\centering
\tiny{\textsf{$\leftarrow$ Unbiased data set}}\\
\tiny{\textsf{Biased data set $\rightarrow$}}\\
\tiny{\textsf{Compared subset:}}\\
\includegraphics[height=0.75em]{aggexplain/legend_all}
}
}
\raisebox{-0.5\height}{\includegraphics[width=0.4\linewidth]{aggexplain/h_all}}
~\\
\noindent\rule{\linewidth}{0.4pt}
~\\
\raisebox{-0.5\height}{\includegraphics[width=0.4\linewidth]{aggexplain/h_label_ok}}
\raisebox{-0.2\height}{
\parbox{0.16\linewidth}{
\centering
\tiny{\textsf{$\leftarrow$ Unbiased data set}}\\
\tiny{\textsf{Biased data set $\rightarrow$}}\\
\tiny{\textsf{Compared subset:}}\\
\includegraphics[height=0.75em]{aggexplain/legend_label}
}
}
\raisebox{-0.5\height}{\includegraphics[width=0.4\linewidth]{aggexplain/h_label_bad}}
~\\
\noindent\rule{\linewidth}{0.4pt}
~\\
\raisebox{-0.5\height}{\includegraphics[width=0.4\linewidth]{aggexplain/h_pred_ok}}
\raisebox{-0.2\height}{
\parbox{0.16\linewidth}{
\centering
\tiny{\textsf{$\leftarrow$ Unbiased data set}}\\
\tiny{\textsf{Biased data set $\rightarrow$}}\\
\tiny{\textsf{Compared subset:}}\\
\includegraphics[height=0.75em]{aggexplain/legend_pred}
}
}
\raisebox{-0.5\height}{\includegraphics[width=0.4\linewidth]{aggexplain/h_pred_bad}}
~\\
\noindent\rule{\linewidth}{0.4pt}
~\\
\raisebox{-0.5\height}{\includegraphics[width=0.4\linewidth]{aggexplain/h_correct_ok}}
\raisebox{-0.2\height}{
\parbox{0.16\linewidth}{
\centering
\tiny{\textsf{$\leftarrow$ Unbiased data set}}\\
\tiny{\textsf{Biased data set $\rightarrow$}}\\
\tiny{\textsf{Compared subset:}}\\
\includegraphics[height=0.75em]{aggexplain/legend_correct}
}
}
\raisebox{-0.5\height}{\includegraphics[width=0.4\linewidth]{aggexplain/h_correct_bad}}
\caption[Comparison of different subset selections on both models.]{
Comparison of different subset selections on both the unbiased model (left side) and the biased model (right side).
Features are sorted per segment by most important at the top-left, row-wise to least important at the bottom-right.
Note, the feature ``Living Area'', in both the ``Label'' (\ie, ground truth) and ``Pred.'' (\ie, model prediction), has flipped outcomes for the biased model (right side).
Each subset has a different color palette to not confuse different selections with each other.
}
\label{figs:good_vs_bad}
\end{figure}

### 2018-04-03 aggexplain/introduction.tex
fix figures and some citations

@@

\begin{figure*}[ht!]
\centering
\includegraphics[width=0.7\linewidth]{aggexplain/full_histogram}
\caption{
The full interface illustrating the aggregated histogram view.
The user is comparing the model's prediction of ``high'' house prices (orange) to the prediction of ``low'' prices (purple).
The user hovers over the feature ``House Style'' revealing a more detailed description, whether the feature is categorical or numeric, and the importance / feature weight for each of the subsets.
}
\label{figs:full_hist_view}
\end{figure*}

>>>>

\begin{figure}[t]
 \centering
 \includegraphics[width=\linewidth]{aggexplain/teaser.pdf}
 \caption[Showing the four study conditions.]{
Showing the four conditions of our study: (1) instance-level explanations; (2) only instances; (3) only aggregated features; (4) aggregated features with explanations.
On the left and the top are consistent parts of the user interface showing: (a) the problem description; (b) the confusion matrix; (c) the subset selector.
%\adam{This teaser is a bit confusing. Would it be better to just have the complete UI here and this figure later.}
% I keep it for now
 }
 \label{figs:teaser}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{aggexplain/full_histogram}
\caption[The full interface illustrating the aggregated histogram view.]{
The full interface illustrating the aggregated histogram view.
The user is comparing the model's prediction of ``high'' house prices (orange) to the prediction of ``low'' prices (purple).
The user hovers over the feature ``House Style'' revealing a more detailed description, whether the feature is categorical or numeric, and the importance / feature weight for each of the subsets.
}
\label{figs:full_hist_view}
\end{figure}

### 2018-04-03 aggexplain/motivation.tex
fix figures and some citations

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{aggexplain/histogram}
\caption{
Comparing the distribution of values by ``Actual Label'' (\ie, ground truth).
The height of the bars show the percentage of values within the respective subset (green for ``high'' outcomes and pink for ``low'' outcomes).
The average feature weight of each subset is shown next to the feature name.
This is only visible in the condition including explanations.
}
\label{figs:histogram}
\end{figure}

@@

\begin{figure*}[ht!]
\centering
\includegraphics[width=0.7\linewidth]{aggexplain/full_table_expl}
\caption{
The full interface illustrating the table view showing individual instances.
The user is comparing the model's prediction of ``high'' house prices (orange) to the prediction of ``low'' prices (purple).
The feature ``Living Area'' is ordered by ascending values and the user hovers over the cell with the value ``605''.
}
\label{figs:full_view}
\end{figure*}

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=0.7\linewidth]{aggexplain/histogram}
\caption[Comparing the distribution of values by ``Actual Label''.]{
Comparing the distribution of values by ``Actual Label'' (\ie, ground truth).
The height of the bars show the percentage of values within the respective subset (green for ``high'' outcomes and pink for ``low'' outcomes).
The average feature weight of each subset is shown next to the feature name.
This is only visible in the condition including explanations.
}
\label{figs:histogram}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{aggexplain/full_table_expl}
\caption[The full interface illustrating the table view.]{
The full interface illustrating the table view showing individual instances.
The user is comparing the model's prediction of ``high'' house prices (orange) to the prediction of ``low'' prices (purple).
The feature ``Living Area'' is ordered by ascending values and the user hovers over the cell with the value ``605''.
}
\label{figs:full_view}
\end{figure}

### 2018-04-03 aggexplain/relatedwork.tex
fix figures and some citations

\subsection{Effectiveness of instance-level explanations}
When introducing their algorithm, LIME (Ribeiro~\etal~\cite{DBLP:journals/corr/RibeiroSG16, anchors:aaai18},2016arXiv161105817T), the authors conducted experiments to show the effectiveness of their method.
However, instance-level explanations were only inspected individually and not in aggregate form.

>>>>

\subsection{Effectiveness of instance-level explanations}
When introducing their algorithm, LIME (Ribeiro~\etal~\cite{DBLP:journals/corr/RibeiroSG16, anchors:aaai18,2016arXiv161105817T}), the authors conducted experiments to show the effectiveness of their method.
However, instance-level explanations were only inspected individually and not in aggregate form.

### 2018-04-03 aggexplain/results.tex
fix figures and some citations

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{aggexplain/correctness}
\caption{
Comparing whether participants correctly identified the bias in the data, determined from plain-text answers. Note, that adding explanations to the table view hurt performance, whereas adding explanations to the histogram view improved performance.
}
\label{figs:correctness}
\end{figure}

@@

\begin{figure}[b]
\centering
\includegraphics[width=0.4\linewidth]{aggexplain/table_interaction}
\caption{
Interaction with the table view measured by counting how many table cells were hovered by the mouse.
The plot shows the bootstrapped mean and confidence interval for each setting.
}
\label{figs:table_interaction}
\end{figure}

@@

\begin{figure}[b]
\centering
\includegraphics[width=0.6\linewidth]{aggexplain/timing2}
\caption{
Overall completion time of the study divided by condition.
The plot shows the bootstrapped mean and confidence interval for each setting.
There is no significant difference between the conditions.
}
\label{figs:timing}
\end{figure}

@@

% \begin{figure*}
% \centering
% \includegraphics[width=0.75\linewidth]{figs/trust}
% \caption{
% The participants' answer to whether they trust the model with higher accuracy, lower accuracy, or both equally. The orange colored part indicates the proportion of participants that correctly found the bias in the data.
% Note, that some participants valued accuracy more than semantically correct data.
% }
% \label{figs:trust}
% \end{figure*}

>>>>

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{aggexplain/correctness}
\caption[Comparing whether participants correctly identified the bias.]{
Comparing whether participants correctly identified the bias in the data, determined from plain-text answers. Note, that adding explanations to the table view hurt performance, whereas adding explanations to the histogram view improved performance.
}
\label{figs:correctness}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.25\linewidth]{aggexplain/table_interaction}
\caption[Interactions with the table view.]{
Interactions with the table view measured by counting how many table cells were hovered by the mouse.
The plot shows the bootstrapped mean and confidence interval for each setting.
}
\label{figs:table_interaction}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.4\linewidth]{aggexplain/timing2}
\caption[Overall completion time of the study divided by condition.]{
Overall completion time of the study divided by condition.
The plot shows the bootstrapped mean and confidence interval for each setting.
There is no significant difference between the conditions.
}
\label{figs:timing}
\end{figure}

@@

% \begin{figure}
% \centering
% \includegraphics[width=0.75\linewidth]{figs/trust}
% \caption{
% The participants' answer to whether they trust the model with higher accuracy, lower accuracy, or both equally. The orange colored part indicates the proportion of participants that correctly found the bias in the data.
% Note, that some participants valued accuracy more than semantically correct data.
% }
% \label{figs:trust}
% \end{figure}

### 2018-04-03 explainer/inspect.tex
fix figures and some citations

\begin{figure}[b!]
\centering
\includegraphics[width=0.9\linewidth]{explainer/inspect1}
% \begin{subfigure}[b]{0.7\linewidth}
%     \centering
%     \includegraphics[width=\linewidth]{fig/inspect}
%     \caption{~Inspection of ``Diatrizoate Meglumine"}
%     \label{figs:inspect}
% \end{subfigure}
% \hspace{-3em}
% \begin{subfigure}[b]{0.35\linewidth}
%     \centering
%     \includegraphics[width=1.03\linewidth]{fig/sodchl} % warning here is necessary
%     \caption{~and ``Sodium Chloride".}
%     \label{figs:sodchl}
% \end{subfigure}
\caption{
The \textbf{\tabC} showing a matrix of data items as rows and features as columns for the explanations \emph{Diatrizoate Meglumine} and \emph{Sodium Chloride} in the initial data set of the case study (Section~\ref{sec:case_study}).
Rows group identical instances together and show the count on the left side.
Features are sorted by ``relative feature importance" showing from left to right how labels can be separated.
% \aritra{This figure needs annotations: Items, Features, Predicted Label, and what the bars mean}
% \joschi{too much info here}
}
\label{figs:inspect_all}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=\linewidth]{explainer/inspect1}
% \begin{subfigure}[b]{0.7\linewidth}
%     \centering
%     \includegraphics[width=\linewidth]{fig/inspect}
%     \caption{~Inspection of ``Diatrizoate Meglumine"}
%     \label{figs:inspect}
% \end{subfigure}
% \hspace{-3em}
% \begin{subfigure}[b]{0.35\linewidth}
%     \centering
%     \includegraphics[width=1.03\linewidth]{fig/sodchl} % warning here is necessary
%     \caption{~and ``Sodium Chloride".}
%     \label{figs:sodchl}
% \end{subfigure}
\caption[The \tabC.]{
The \textbf{\tabC} showing a matrix of data items as rows and features as columns for the explanations \emph{Diatrizoate Meglumine} and \emph{Sodium Chloride} in the initial data set of the case study (Section~\ref{sec:case_study}).
Rows group identical instances together and show the count on the left side.
Features are sorted by ``relative feature importance" showing from left to right how labels can be separated.
% \aritra{This figure needs annotations: Items, Features, Predicted Label, and what the bars mean}
% \joschi{too much info here}
}
\label{figs:inspect_all}
\end{figure}

### 2018-04-03 explainer/odds_ratio.tex
fix figures and some citations

\begin{figure}
\centering
\begin{subfigure}[b]{\linewidth}
    \hfill
    \includegraphics[height=7em]{explainer/expl_10_size}
    \caption{~Ordered by ``total" size showing the most common explanations.}
    \label{figs:expl_10_size}
\end{subfigure}
\\
\begin{subfigure}[b]{\linewidth}
    \hfill
    \includegraphics[height=7em]{explainer/expl_10_or_pos}
    \caption{~Ordered by ``odds ratio" showing significantly positive explanations.}
    \label{figs:expl_10_or_pos}
\end{subfigure}%
\\
\begin{subfigure}[b]{\linewidth}
    \hfill
    \includegraphics[height=7em]{explainer/expl_10_or_neg}
    \caption{~Ordered by reverse ``odds ratio" showing significantly negative explanations.}
    \label{figs:expl_10_or_neg}
\end{subfigure}%
\\
\begin{subfigure}[b]{\linewidth}
    \hfill
    \includegraphics[height=7em]{explainer/expl_new_uncertain}
    \caption{~Ordered by ``uncertainty" showing item subsets whose predictions are not significant.}
    \label{figs:expl_10_uncertain}
\end{subfigure}
\caption{
Showing different orders in the \textbf{\tabB} for addressing the goals (G2 \& G3) in the case study (Section~\ref{sec:case_study}). The initial dataset is filtered for explanations with $> 20$ data items.
% The \textbf{\tabB} showing the initial case study dataset (see Section~\ref{sec:case_study}) filtered for explanations with $> 20$ data items.
}
\end{figure}

>>>>

\begin{figure}
\centering
\begin{subfigure}[b]{\linewidth}
    \hfill
    \includegraphics[height=7em]{explainer/expl_10_size}
    \caption{~Ordered by ``total" size showing the most common explanations.}
    \label{figs:expl_10_size}
\end{subfigure}
\\
\begin{subfigure}[b]{\linewidth}
    \hfill
    \includegraphics[height=7em]{explainer/expl_10_or_pos}
    \caption{~Ordered by ``odds ratio" showing significantly positive explanations.}
    \label{figs:expl_10_or_pos}
\end{subfigure}%
\\
\begin{subfigure}[b]{\linewidth}
    \hfill
    \includegraphics[height=7em]{explainer/expl_10_or_neg}
    \caption{~Ordered by reverse ``odds ratio" showing significantly negative explanations.}
    \label{figs:expl_10_or_neg}
\end{subfigure}%
\\
\begin{subfigure}[b]{\linewidth}
    \hfill
    \includegraphics[height=7em]{explainer/expl_new_uncertain}
    \caption{~Ordered by ``uncertainty" showing item subsets whose predictions are not significant.}
    \label{figs:expl_10_uncertain}
\end{subfigure}
\caption[Showing different orders in the \tabB.]{
Showing different orders in the \textbf{\tabB} for addressing the goals (G2 \& G3) in the case study (Section~\ref{sec:case_study}). The initial dataset is filtered for explanations with $> 20$ data items.
% The \textbf{\tabB} showing the initial case study dataset (see Section~\ref{sec:case_study}) filtered for explanations with $> 20$ data items.
}
\end{figure}

### 2018-04-03 explainer/workflow.tex
fix figures and some citations

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{explainer/workflow5}
\vspace{-3mm}
\caption{
Our proposed \textbf{Model Diagnostics} workflow extends the conventional \textit{Model Building} workflow in machine learning for enabling domain experts to reason about the semantic validity of the decisions made by any model through multiple linked visualizations of statistical performance summaries, explanations, and item-level distribution of features. By iterating through explanation-level summaries and item-level details, experts are able to generate diagnostic insights about the quality of both the data and the model. This ultimately helps to improve data acquisition and model generation processes belonging to the original workflow.
}
\vspace{-5mm}
\label{figs:workflow}
\end{figure*}

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{explainer/workflow5}
\caption[The Model Diagnostics workflow.]{
Our proposed \textbf{Model Diagnostics} workflow extends the conventional \textit{Model Building} workflow in machine learning for enabling domain experts to reason about the semantic validity of the decisions made by any model through multiple linked visualizations of statistical performance summaries, explanations, and item-level distribution of features. By iterating through explanation-level summaries and item-level details, experts are able to generate diagnostic insights about the quality of both the data and the model. This ultimately helps to improve data acquisition and model generation processes belonging to the original workflow.
}
\label{figs:workflow}
\end{figure}

### 2018-04-03 infuse/introduction.tex
fix figures and some citations

% !TEX root = ../featureselection.tex
\begin{figure*}[ht]
\centering
\includegraphics[width=\linewidth]{infuse/pipeline}
\caption[Steps of a typical predictive modeling pipeline.]{Steps of a typical predictive modeling pipeline.  For each step, we provide the details of the running example we use throughout the paper.
}
\label{fig:pipeline}
\end{figure*}

The visualization research community has usually focused on developing techniques and systems to support the analysis of datasets, with limited analysis of the relationship between datasets and the construction of models on top of them. However, there are a growing number of data scientists interested in more than just interpreting their data: they want to understand their data and predictive probabilities associated with them. Providing visual support for this kind of task has become important as many existing applications on the market and in scientific settings need to solve problems that are predictive in nature, e.g. prediction of customer behavior, diseases, drug effectiveness.

@@

@@

@@

\begin{figure*}[t]
\centering
\includegraphics[width=\linewidth]{infuse/system}
\caption[Breakdown of the user interface of \infuse.]{
An overview of \infuse, a system for interactive feature selection.  On the left, the Feature View provides a way to visualize an overview of all features grouped by type and then sorted by importance.  The color key for the feature types and subtypes are shown at the bottom.  The buttons and combo boxes at the bottom can be used to switch layouts and define the axes of the scatterplot view shown in
Figure~\ref{fig:scatter}.  On the top-right, the List View provides a sorted list of all features, useful for selections.
This list can be filtered using the search box above. Currently only features containing the term ``gl" are shown.
The remaining features are sorted by the number and position
of the search term occurrences.
On the bottom-right, the Classifier View (Figure~\ref{fig:classifier}) provides access to the quality scores of each model.  Users can also select features and build custom models with the Interactive Model Builder.
}
\label{fig:system}
\end{figure*}

>>>>

% !TEX root = ../featureselection.tex
The visualization research community has usually focused on developing techniques and systems to support the analysis of datasets, with limited analysis of the relationship between datasets and the construction of models on top of them. However, there are a growing number of data scientists interested in more than just interpreting their data: they want to understand their data and predictive probabilities associated with them. Providing visual support for this kind of task has become important as many existing applications on the market and in scientific settings need to solve problems that are predictive in nature, e.g. prediction of customer behavior, diseases, drug effectiveness.

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{infuse/teaser}
\caption[Overview of \infuse.]{
An overview of \infuse, a visual analytics tool that supports users to understand the predictive power of features in their models.  Each feature is ranked by various feature selection algorithms, and the ranking information is visualized in each of the three views within the system.  On the left, the Feature View provides a way to visualize an overview of all features according to their rank using a variety of layouts.  On the top-right, the List View provides a sorted list of all features, useful for selections.  On the bottom-right, the Classifier View provides access to the quality scores of each model.  Each of the views are coordinated, and users can brush between all three views.
}
\label{fig:teaser}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{infuse/pipeline}
\caption[Steps of a typical predictive modeling pipeline.]{Steps of a typical predictive modeling pipeline.  For each step, we provide the details of the running example we use throughout the paper.
}
\label{fig:pipeline}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{infuse/system}
\caption[Breakdown of the user interface of \infuse.]{
An overview of \infuse, a system for interactive feature selection.  On the left, the Feature View provides a way to visualize an overview of all features grouped by type and then sorted by importance.  The color key for the feature types and subtypes are shown at the bottom.  The buttons and combo boxes at the bottom can be used to switch layouts and define the axes of the scatterplot view shown in
Figure~\ref{fig:scatter}.  On the top-right, the List View provides a sorted list of all features, useful for selections.
This list can be filtered using the search box above. Currently only features containing the term ``gl" are shown.
The remaining features are sorted by the number and position
of the search term occurrences.
On the bottom-right, the Classifier View (Figure~\ref{fig:classifier}) provides access to the quality scores of each model.  Users can also select features and build custom models with the Interactive Model Builder.
}
\label{fig:system}
\end{figure}
